The shape of dark matter halos in the AURIGA simulations

### 2019-03-05 Paper/paper.tex
edit

\subsection{Comparison with observational constraints}

\begin{table*}
\begin{tabular}{|l|cc|cc|c|c|p{4cm}|}\hline
Reference&$q_{\rho}$&$s_{\rho}$&$q_{\phi}$&$s_{\phi}$&$R$&$\theta$&comment\\ \hline \hline
\citet{Olling_and_Merrifield_2000}& $\mathbf{1.00}$ & $\mathbf{0.80}$ & & & $\simeq 8$kpc & $0^{\circ}$&Method: Stellar dynamics and HI density. \\\hline
\citet{Law_and_Majewski_2009}&&&$\mathbf{0.83}$&$\mathbf{0.67}$& $\lesssim 60$kpc&$90^{\circ}$&Mid-axis orientation. Method: Sagittarius stream\\\hline
%
\citet{Law_and_Majewski_2010}&&&$\mathbf{0.99}$&$\mathbf{0.72}$& $[20$kpc,$60$kpc$]$&$90^{\circ}$&Mid-axis orientation, Method: Sagittarius stream\\\hline
%
\citet{Loebman_et_al._2012}&$\mathbf{1.00}$&$\mathbf{0.47}$&&&$\sim 20$kpc &$0^{\circ}$&Method: SDSS statistics\\\hline
%
\citet{Deg_and_Widrow_2013}&$0.72$&$0.28$&$0.82$&$0.40$&$[20$kpc,$60$kpc$]$&$90^{\circ}$& Mid-axis orientation. Method: Sagittarius stream\\\hline
%
\multirow{2}{*}{\citet{Vera-Ciro_and_Helmi_2013}}&&&$\mathbf{1.00}$&$\mathbf{0.90}$&$\lesssim 10$kpc&$0^{\circ}$ & Method: Sagittarius stream \& LMC \\
&&&$\mathbf{0.90}$&$\mathbf{0.80}$&$\gtrsim 10$kpc&$90^{\circ}$& Mid-axis orientation on the outside. \\\hline
%
\citet{Bovy_et_el._2016}&$\mathbf{0.95}$&$\mathbf{0.95}$&&&$\lesssim 20$kpc&$90^{\circ}$ & Method: Stellar streams\\\hline
%
\citet{Bowden_et_al._2016}&&&$\mathbf{[0.5,0.66]}$&$\mathbf{[0.5,0.66]}$&$[5$kpc,$10$kpc$]$&$90^{\circ}$& Weak constraint on prolate halo. Method: SDSS stars dynamics.\\\hline
%
\multirow{2}{*}{\citet{Banerjee_and_Chanda_2011}}&$\mathbf{1}$&$\mathbf{1}$&&&$9$kpc&$0^{\circ}$&Method: HI gas. \\
&$\mathbf{0.5}$&$\mathbf{0.5}$&&&$24$kpc&$0^{\circ}$&Monotonical change between radial regimes.\\\hline
%
\cite{Johnston_et_al._2005}&$\mathbf{1}$&$\mathbf{[0.83-0.92]}$&&&$\lesssim 60$kpc&$0^{\circ}$&Method: Sagittarius stream\\\hline\hline
\end{tabular}
\caption{(TODO: compute analogues in isopotential or isodensity according to Binney and Tremaine)}
\end{table*}
%\end{multicols}

\begin{table*}
%\setlength{\tabcolsep}{3pt}
\begin{tabular}{|l|cc|cc|c|p{4cm}|}\hline
Reference&$q_{\rho}$&$s_{\rho}$&$q_{\phi}$&$s_{\phi}$&$R$&comment\\ \hline \hline
\citet{Chua_et_al._2018}&$\mathbf{0.88\pm0.10}$&$\mathbf{0.70\pm0.11}$&&&$0.15R_{200}$& Illustris\\\hline
%
\citet{Bryan_et_al._2013}&$\mathbf{[0.84,0.86]}$&$\mathbf{[0.66,0.70]}$&&&$R_{200}$& For different cosmologies and feedback recipies. Calculated from a fit at $M_\odot=10^12$\\\hline
%
\multirow{2}{*}{\citet{Abadi_et_al._2010}}&&&$\mathbf{0.98}$&$\mathbf{0.85}$&-& Almost independent of radius. No feedback: boundary case\\\hline
\end{tabular}
\caption{(TODO: compute analogues in isopotential or isodensity according to Binney and Tremaine)}
\end{table*}

To be able to compare our results with observational values, we must relate the calculated quanties with their corresponding isopotential analogue, in which observational constraints are usually presented. For this purpose, we run a simple algorithm to find an approximation of the shape of the isopotential contour. Here, we calculate the mean and standard deviation of the potential over a spherical shell of width equals to $10\%$ of the radius at which it is sampled. Then, we calculate the inertia tensor of particles with potential within $1\sigma$ around the mean potential and calculate its triaxial characterization with the reduced inertia tensor. We repeat the process of calculating the potential mean and standard deviation until convergence is achieved with tolerance of $10^{-4}$. We repeat this process for the different radii from table \ref{tabe:isopotential}. \\

@@

As a check of consistency, we compare our new isopotential shape
results with the analytic expression of the
$(1-q_{\phi})\frac{1}{3}(1-q_{\rho})$ \citep{Binney_and_Tremaine_2008}, taking the volume-enclosed axial ratios
as an approximation for the isodensity contour ratios
$q_{\rho}$. Although this analytic expression if meant to be used for
logarithmic axisymmetric halos, it works well as a first approximation
for nearly axisymmetric halos as those produced by our disced
galaxies. We find that the difference between the real and the
analytic isopotential axial ratios is not bigger than $quantity
percent$. With this, we may now present on table
\ref{tabe:isopotential} our observationally-comparable results for MHD
halos at two important radii one corresponding to the approximate
regimes where the MW DM halo is usually constrained.

>>>>

\section{Analysis}

\subsection{Comparison against observational constraints}

\begin{table*}
\begin{tabular}{|l|cc|c|c|p{7cm}|}\hline
Reference&$q_{\phi}$&$s_{\phi}$&$R$&$\theta$&
Methodology\\ \hline \hline

\citet{Bowden_et_al._2016}& 0.5-0.66 &0.5-0.66& 5kpc - 10 kpc& $90^{\circ}$&
Weak constraint on prolate halo. SDSS stars dynamics.\\\hline

\multirow{2}{*}{\citet{Vera-Ciro_and_Helmi_2013}}&${1.00}$&${0.90}$&$\lesssim
10$kpc&$0^{\circ}$ & Sagittarius stream \& LMC \\
&${0.90}$&${0.80}$&$\gtrsim 10$kpc&$90^{\circ}$&
Mid-axis orientation on the outside. \\\hline

\citet{Law_and_Majewski_2009}&${0.83}$&${0.67}$&
$\lesssim 60$kpc&$90^{\circ}$&Mid-axis orientation. Sagittarius
stream\\\hline

\citet{Law_and_Majewski_2010}&${0.99}$&${0.72}$&
$20$kpc-$60$kpc$$&$90^{\circ}$&Mid-axis orientation, Sagittarius
stream\\\hline

\citet{Deg_and_Widrow_2013} & $0.82$&$0.40$&
$20$kpc-$60$kpc$ $&$90^{\circ}$&
Mid-axis orientation. Sagittarius stream\\\hline

\end{tabular}
\caption{{\bf Cuales son las incertidumbres? Con respecto a que eje
    esta medido theta?}}
\end{table*}
%\end{multicols}

\begin{table*}
\begin{tabular}{|l|cc|c|c|p{7cm}|}\hline
Reference&$q_{\rho}$&$s_{\rho}$&$R$&$\theta$&
Methodology\\ 
\hline \hline \citet{Olling_and_Merrifield_2000}& $1.00$
& $0.80$ &  $\simeq 8$kpc & $0^{\circ}$&Stellar dynamics and HI
density. \\\hline

\multirow{2}{*}{\citet{Banerjee_and_Chanda_2011}}&${1}$&${1}$&$9$kpc&$0^{\circ}$&Method:
HI
gas. \\ &${0.5}$&${0.5}$&$24$kpc&$0^{\circ}$&Monotonical
change between radial regimes.\\\hline

\citet{Loebman_et_al._2012}&${1.00}$&${0.47}$&$\sim
20$kpc &$0^{\circ}$&Method: SDSS statistics\\\hline

\cite{Johnston_et_al._2005}&${1}$&$0.83-0.92$&$\lesssim
60$kpc&$0^{\circ}$&Method: Sagittarius stream\\\hline

\citet{Bovy_et_el._2016}&${0.95}$&${0.95}$&$\lesssim
20$kpc&$90^{\circ}$ & Stellar streams\\\hline

\citet{Deg_and_Widrow_2013}&$0.72$&$0.28$&
$20$kpc-$60$kpc$ $&$90^{\circ}$&
Mid-axis orientation. Sagittarius stream\\\hline

\citet{Abadi_et_al._2010} &${0.98}$&${0.85}$& & & Simulations. Almost
  independent of radius. No feedback: boundary case\\\hline

\end{tabular}
\caption{{\bf Cuales son las incertidumbres? Con respecto a que eje
    esta medido theta?}}
\end{table*}

\begin{table*}
\begin{tabular}{|l|cc|cc|c|p{4cm}|}\hline
Reference&$q_{\rho}$&$s_{\rho}$&$q_{\phi}$&$s_{\phi}$&$R$& Methodology
\\ \hline \hline
\citet{Chua_et_al._2018}&${0.88\pm0.10}$&${0.70\pm0.11}$&&&$0.15R_{200}$& Illustris\\\hline

\citet{Bryan_et_al._2013}&$0.84-0.86$&$0.66-0.70$&&&$R_{200}$&
For different cosmologies and feedback recipies. Calculated from a fit
at $M_\odot=10^12$\\\hline

\multirow{2}{*}{\citet{Abadi_et_al._2010}}&&&${0.98}$&${0.85}$&-&
Almost independent of radius. No feedback: boundary case\\\hline 
\end{tabular}
\caption{(TODO: compute analogues in isopotential or isodensity
  according to Binney and Tremaine) {\bf Al final estos numeros son
    isodensidad o isopotencial}}
\end{table*}

To compare our results against observational values have we to
translate first the isodensity into the corresponding isopotential
rations, in which observational constraints are usually presented.
For this purpose, we run a simple algorithm to find an
approximation of the shape of the isopotential contour.

First, we calculate the mean and standard deviation of the potential over a
spherical shell of width equals to $10\%$ of the radius at which it is
sampled. 
Then, we calculate the inertia tensor of particles with potential
within $1\sigma$ around the mean potential and calculate its triaxial
characterization with the reduced inertia tensor. 
We repeat the
process of calculating the potential mean and standard deviation until
convergence is achieved with tolerance of $10^{-4}$.

{\bf Aqui hace falta una grafica comparando explicitamente los
  valores para q y para s, esto para diferentes radios.
  Otro punto importante es que se tiene q, pero como se
  relaciona s?
  Ademas hay que calcularlos para los radius aproximados que se
  presentan en las tablas observacionales y en las de simulaciones.
  10-20kpc se acerca a 1/8Rvir, 20-25 1/4Rvir, 60kpc aproximadamente 1/2Rvir
}

As a consistency check we compare the isopotential results with the analytic expectation of the
$(1-q_{\phi})\approx \frac{1}{3}(1-q_{\rho})$
\citep{Binney_and_Tremaine_2008}, taking the volume-enclosed axial
ratios  as an approximation for the isodensity contour ratios
$q_{\rho}$. 
Although this analytic expression if meant to be used on the outer
regions of logarithmic axisymmetric halos, it works well as a first
approximation for nearly axisymmetric halos as those produced by our
simulations 
We find that the difference between the measured and the
approximated isopotential axial ratios is not bigger than $quantity
percent$ {\bf }.

We repeat this process for the different radii from table

@@
